---
title: "STAT293 Code"
author: "Nancy Lopez"
date: "`r Sys.Date()`"
output: pdf_document
---

# Step 1: Packages

```{r}
# install.packages("multivar")
# install.packages("glmnet")
# install.packages("Matrix")

library(glmnet)
library(Matrix)
library(multivar)
library(ggplot2)
```

# Step 2: Simulation Paramaters 

```{r}

K <- 3 # the number of individials  
T <- 12 # the number of time points 
d <- 4 # the number of variables
p_total <- 0.05 # the proportion of nonzeros in the common mu
heterogeneity <- "medium" 
sigma_ep <- 0.1 # the noise level 
holdout <- 3 

```

# Step 3: Helper Function 

randomly selects a position in a dxd matrix 

```{r}

pick_positions <- function(d, n_nonzero) {
  total <- d * d
  idx <- sample(1:total, n_nonzero, replace = FALSE)
  return(idx)
}
```

# Step 4: Heterogeneity 

```{r}

frac_shared <- switch(heterogeneity,
                      "low"    = 2/3,
                      "medium" = 1/2,
                      "high"   = 1/3)

frac_unique <- 1 - frac_shared

```

# Step 5: Shared Matrix

```{r}

# total expected nonzeros at 0.05
n_nonzero_total <- max(1, round(p_total * d * d))

# number of shared 
n_shared <- max(1, round(frac_shared * n_nonzero_total))

# pick shared positions
shared_positions <- pick_positions(d, n_shared)

# making the shared matrix  
M <- matrix(0, d, d)
M_vals <- runif(n_shared, 0.1, 0.9) * sample(c(-1,1), n_shared, replace=TRUE)
M[shared_positions] <- M_vals

```

# Step 6: Individual Matrix 

```{r}

individual_B <- list()

for (k in 1:K) {
  n_unique <- max(1, round(frac_unique * n_nonzero_total))

  #they do NOT overlap with shared positions
  possible_positions <- setdiff(1:(d*d), shared_positions)
  unique_positions <- sample(possible_positions, n_unique, replace = FALSE)

  # \delta_k
  Rk <- matrix(0, d, d)
  R_vals <- runif(n_unique, 0.1, 0.9) * sample(c(-1,1), n_unique, replace=TRUE)
  Rk[unique_positions] <- R_vals

  # final individual B_k
  B_k <- M + Rk

  # stability 
  attempt <- 1
  while (max(abs(eigen(B_k)$values)) >= 1) {
    R_vals <- runif(n_unique, 0.1, 0.9) * sample(c(-1,1), n_unique, replace=TRUE)
    Rk[unique_positions] <- R_vals
    B_k <- M + Rk
    attempt <- attempt + 1
    if (attempt > 50) {
      warning("Could not get stable matrix after 50 tries.")
      break
    }
  }
  individual_B[[k]] <- B_k
}

```

# Step 7: Generate ILD for each ind 

```{r}

generate_VAR <- function(B, T, d, sigma) {
  X <- matrix(0, T, d)
  X[1, ] <- rnorm(d)

  for (t in 2:T) {
    X[t, ] <- B %*% X[t-1, ] + rnorm(d, 0, sigma)
  }
  return(X)
}

data_list <- lapply(individual_B,
                    function(B) generate_VAR(B, T, d, sigma_ep))

```

# Step 8: Training & Testing

```{r}

train_data <- lapply(data_list, function(X) X[1:(T - holdout), ])
test_data  <- lapply(data_list, function(X) X[(T - holdout + 1):T, ])

```

# Step 9: Individual LASSO

```{r}

fit_lasso_individual <- function(X, lambda_ = NULL) {
  X_lag <- X[-nrow(X), ] # X_{t-1}
  Y <- X[-1, ] # X_t
  
  d <- ncol(X)
  fit_list <- vector("list", d)
  
  for (j in 1:d) {
    fit_list[[j]] <- cv.glmnet(
      X_lag,
      Y[, j],
      alpha = 1, # thee LASSO penalty
      lambda = lambda_ #the glmnet func chooses one if lambda is NULL
      )
    }
  return(fit_list)
}

```

# Step 10: Get Estimated beta matrix 

```{r}

B_from_lasso <- function(fit_list, d) {
  # this converts list of cv.glmnet fits into a d x d matrix
  
  B_hat <- matrix(0, d, d)

  for (j in 1:d) {
    coefs <- as.vector(coef(fit_list[[j]], s = "lambda.min"))
    B_hat[j, ] <- coefs[-1] # removes the intercept
  }

  return(B_hat)
}


```

# Step 11: Fit LASSO for all individuals 

```{r}

# fit LASSO for each person in train_data
lasso_fits_all <- lapply(train_data, fit_lasso_individual)

# Convert all fits into estimated B matrices
B_hat_all <- lapply(lasso_fits_all, function(fit_list) {
  B_from_lasso(fit_list, d)
})

```


# Step 12: Compare LASSO B vs true B

```{r}
cat("Estimated B for individual 1:\n")
print(round(B_hat_all[[1]], 3))

cat("\nTrue B for individual 1:\n")
print(round(individual_B[[1]], 3))
```

# Step 13: Multi-VAR 

```{r}
multi_var_model <- constructModel(
  data = train_data,
  lag = 1,
  weightest   = "lasso",
  lassotype = "standard", # standard multi-VAR (not adaptive)
  cv = "rolling",
  window = 3, # must be >1 for real rolling CV
  standardize = TRUE,
  lambda1 = 10^seq(-1, 2, length.out = 10),
  lambda2 = 10^seq(-1, 2, length.out = 10)
)

cv_results <- cv.multivar(multi_var_model)

```

Step 14: Extract beta matrices for multivar

```{r}
B_hat_multi_1 <- cv_results$mats[[1]]
round(B_hat_multi_1, 3)

```

# Step 15: Compare Multi-VAR beta to true beta

```{r}
cat("Multi-VAR estimated B for individual 1:\n")
print(round(B_hat_multi_1, 3))

cat("\nTrue B for individual 1:\n")
print(round(individual_B[[1]], 3))

```

# Step 16: Multi-VAR error

```{r}
error_multi <- sum((B_hat_multi_1 - individual_B[[1]])^2)
error_multi

```

# Step 17: Adaptive Multi-VAR

```{r}
multi_var_adaptive <- constructModel(
  data = train_data,
  lag = 1,
  weightest = "lasso",
  lassotype = "adaptive",
  cv = "rolling",
  window = 3,
  standardize = TRUE,
  lamadapt = TRUE,
  lambda1 = 10^seq(-1, 2, length.out = 10),
  lambda2 = 10^seq(-1, 2, length.out = 10)
)

cv_adaptive <- cv.multivar(multi_var_adaptive)

```

Step 18: Extract beta matrices for adaptive multivar

```{r}
B_hat_adaptive_1 <- cv_adaptive$mats[[1]]
round(B_hat_adaptive_1, 3)
```

# Step 19: Compare adaptive Multi-VAR beta to true beta

```{r}
cat("Adaptive Multi-VAR estimated B for individual 1:\n")
print(round(B_hat_adaptive_1, 3))

cat("\nTrue B for individual 1:\n")
print(round(individual_B[[1]], 3))

```

# Step 20: adaptive Multi-VAR error

```{r}
error_adaptive <- sum((B_hat_adaptive_1 - individual_B[[1]])^2)
error_adaptive
```

Step 21: Finding Sensitivity and Specificity Function

```{r}
sensitivity_specificity <- function(B_true, B_est, thr) {
  true_nonzero <- abs(B_true) > thr
  est_nonzero <- abs(B_est) > thr
  
  TP <- sum(est_nonzero & true_nonzero)
  FN <- sum(!est_nonzero & true_nonzero)
  TN <- sum(!est_nonzero & !true_nonzero)
  FP <- sum(est_nonzero & !true_nonzero)
  
  sensitivity <- TP / (TP + FN)
  specificity <- TN / (TN + FP)
  
  return(list(sensitivity = sensitivity, specificity = specificity))
  }
```

Step 22: Comparing Sensitivity and Specificity

```{r}
# LASSO
metrics_lasso_1 <- sensitivity_specificity(
  B_true = individual_B[[1]],
  B_est = B_hat_all[[1]],
  thr = 0.03
  )

# multi var
metrics_multi_1 <- sensitivity_specificity(
  B_true = individual_B[[1]],
  B_est = B_hat_multi_1,
  thr = 0.03
  )

# adaptive multi var

metrics_adapt_1 <- sensitivity_specificity(
  B_true = individual_B[[1]],
  B_est = B_hat_adaptive_1,
  thr = 0.03
  )

# results
results_ss <- data.frame(
  Model= c("LASSO", "Multi-VAR", "Adaptive Multi-VAR"),
  Sensitivity = c(metrics_lasso_1$sensitivity,
                  metrics_multi_1$sensitivity,
                  metrics_adapt_1$sensitivity),
  Specificity = c(metrics_lasso_1$specificity,
                  metrics_multi_1$specificity,
                  metrics_adapt_1$specificity)
  )

results_ss


```
Step 23: Comparing RMSFE (Root Mean Square Forecast Error)

```{r}
compute_rmsfe <- function(B_est, test, train_last) {
  preds <- matrix(0, nrow(test), ncol(test))
  x_prev <- train_last
  for (t in 1:nrow(test)) {
    preds[t,] <- B_est %*% x_prev
    x_prev <- test[t,]
    }
  return(sqrt(mean((preds - test)^2)))
}

# LASSO
rmsfe_lasso <- compute_rmsfe(
  B_est = B_hat_all[[1]],
  test = test_data[[1]],
  train_last = train_data[[1]][nrow(train_data[[1]]), ]
  )

# multi var
rmsfe_multi <- compute_rmsfe(
  B_est = B_hat_multi_1,
  test = test_data[[1]],
  train_last = train_data[[1]][nrow(train_data[[1]]), ]
  )

# adaptive multi var

rmsfe_adaptive <- compute_rmsfe(
  B_est = B_hat_adaptive_1,
  test = test_data[[1]],
  train_last = train_data[[1]][nrow(train_data[[1]]), ]
  )

data.frame(
  Model = c("LASSO", "Multi-VAR", "Adaptive Multi-VAR"),
  RMSFE = c(rmsfe_lasso, rmsfe_multi, rmsfe_adaptive)
  )

```
# PLOTS 

```{r}

library(reshape2)

df_long <- melt(results_ss, id.vars = "Model")

ggplot(df_long, aes(x = Model, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Sensitivity and Specificity for Each Method",
       x = "Model",
       y = "Value",
       fill = "") +
  ylim(0, 1) +
  theme_minimal()

rmsfe_df <- data.frame(
  Model = c("LASSO", "Multi-VAR", "Adaptive Multi-VAR"),
  RMSFE = c(rmsfe_lasso, rmsfe_multi, rmsfe_adaptive)
)

ggplot(rmsfe_df, aes(x = Model, y = RMSFE, fill = Model)) +
  geom_bar(stat = "identity") +
  labs(title = "Forecasting Accuracy (RMSFE)",
       x = "Model",
       y = "RMSFE") +
  theme_minimal()

```